---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r loading_data, echo = TRUE}
require(data.table)
require(lubridate)
require(lattice)

activity <- data.table(read.table(unz('~/Github/RepData_PeerAssessment1/activity.zip', 'activity.csv'),
                                  header = TRUE, sep = ','))

activity[, steps := as.numeric(steps)]
```

## What is mean total number of steps taken per day?
```{r mean_steps, echo = TRUE}
day_steps <- activity[, .(aver_steps = sum(steps, na.rm = TRUE)), by = ymd(date)]
hist(day_steps$aver_steps, 20, main = 'Total steps per day', xlab = 'steps', las = 1)
unlist(lapply(list(mean = mean, median = median), function(f) f(day_steps$aver_steps, na.rm = TRUE)))
```

## What is the average daily activity pattern?
```{r daily_pattern, echo = TRUE}
interval_steps <- activity[, .(aver_steps = mean(steps, na.rm = TRUE)), by = interval]
plot(interval_steps, type ='l', las = 1, ylab = 'Number of steps')
interval_steps[which.max(aver_steps)]
```

## Imputing missing values
For each NA we impute the mean number of steps of their correspoding time interval.
```{r missing_values, echo = TRUE}
sum(is.na(activity))
invisible(lapply(interval_steps[, interval], function(x){
        value <- interval_steps[interval == x, aver_steps]
        activity[is.na(steps) & interval == x, steps := value]
}))
day_steps <- activity[, .(aver_steps = sum(steps)), by = ymd(date)]
hist(day_steps$aver_steps, 20, main = 'Total steps per day', xlab = 'steps', las = 1)
unlist(lapply(list(mean = mean, median = median), function(f) f(day_steps$aver_steps)))
```
In fact, the mean and the median are differents from the above result.

## Are there differences in activity patterns between weekdays and weekends?
```{r weekdays_pattern, echo = TRUE}
Sys.setlocale("LC_TIME","English")
activity[, weekdays := factor(ifelse(weekdays(ymd(date)) %in% c('Saturday', 'Sunday'),'weekend', 'weekday'))]
week_steps <- activity[, .(aver_steps = mean(steps, na.rm = TRUE), weekdays), by = interval]
xyplot(aver_steps ~ interval | weekdays, data = week_steps, layout = c(1, 2), type="l", xlab = "Interval", ylab = "Number of steps")
```
